The combination of reporting bias and underpowered study designs has substantially exaggerated the motor learning benefits of self-controlled practice and enhanced expectancies: A meta-analysis

### 2022-09-06 docs/preprint/preprint.tex
update: fix email address; minor cosmetic tweaks

\noindent \textsuperscript{$\dagger$}Co-Senior Authors \newline
\newline
\noindent \addORCIDlink{Brad McKay}{0000-0002-7408-2323} \newline
\noindent \addORCIDlink{Mariane F.B. Bacelar}{0000-0002-0388-1769} \newline
\noindent \addORCIDlink{Juliana O. Parma}{0000-0002-3118-9921} \newline
\noindent \addORCIDlink{Matthew W. Miller}{0000-0002-7896-7217} \newline
\noindent \addORCIDlink{Michael J. Carter}{0000-0002-0675-4271} \newline       
\newline
\noindent Data and code: \url{https://osf.io/vfza7/} \newline
\newline
\noindent \textbf{Corresponding authors:} Brad McKay (\href{mailto:bradmckay8@gmail.com}{\nolinkurl{bradmckay8@gmail.com}}; \href{mailto:mckayb9@mcmaster.ca}{\nolinkurl{mckayb9@mcmaster.ca}}) and Michael J. Carter (\href{mailto:michaelcarter@mcmaster.ca}{\nolinkurl{michaelcarter@mcmaster.ca}}; \href{mailto:motorlab@mcmaster.ca}{\nolinkurl{motorlab@mcmaster.ca}})

@@

A challenge for most reporting bias models is large heterogeneity in true effects. While RoBMA appears to perform well with moderate levels of heterogeneity, its performance has not been evaluated when heterogeneity is high and the performance of each of its constituent reporting bias models suffers with high heterogeneity (Carter et al., 2019). The \emph{z}-curve model was designed specifically to perform well regardless of heterogeneity (Bartoš \& Schimmack, in-press). While \emph{z}-curve does not provide adjusted effect size estimates (and such estimates may be meaningless with high heterogeneity), it instead estimates the average underlying power of included experiments. A significant difference between the estimated power of studies and the observed proportion of significant results can indicate the presence of reporting bias in a literature.

Here, we leveraged state-of-the-art robust Bayesian meta-analysis and \emph{z}-curve methods to re-analyze the meta-analyses by McKay and colleagues (in-press) and Bacelar and colleagues (2022). Considering the potential importance of OPTIMAL theory for the field of motor learning and performance (see T. D. Lee \& Carnahan, 2021 for a discussion), it is imperative that the evidence buttressing its motivation predictions be evaluated as rigorously as possible. Critically, our re-analysis addresses limitations in both previous meta-analyses. First, the analysis of enhanced expectancies fit only one bias correction model---the trim-and-fill method (Duval \& Tweedie, 1998)---and that model has been shown to result in exaggerated effect size estimates and severely inflated Type 1 error rates in the presence of publication bias and small or null effects (Bartoš et al., 2022; Carter et al., 2019). Second, although the results from multiple reporting bias models coalesced around small effect sizes (ranged from \emph{g} = -.11 to \emph{g} = .26) in the analysis of self-controlled practice, there are no principled reasons for preferring one estimate over another. We now consider a wider range of plausible models of reporting bias than that used in the previous meta-analyses. We also leverage Bayesian model averaging to upweight the best performing models, which has the advantage of evaluating single model-averaged posterior distributions for each parameter of interest. Lastly, we fit \emph{z}-curve models to the data in both meta-analyses. With this technique, the average power can be estimated and compared to the rate of significant results, providing crucial insight into the quality of the evidence-base supporting motivation predictions in OPTIMAL theory.

@@

The original meta-analyses followed similar inclusion criteria and data extraction policies (see Figure \ref{fig:fig1}). All data were extracted by two independent researchers with conflicts settled by including a third researcher. The literature search ended in August of 2019 in the self-controlled practice meta-analysis, and 10 months later in June 2020 in the enhanced expectancies meta-analysis.

@@

\caption{\normalfont \small \textbf{Inclusion criteria, dependent variable selection, studies screended and outcomes included in each meta-analysis.}}\label{fig:fig1}
\end{figure}

@@

The second class of models fit in the RoBMA-PSMA method model the relationship between observed effect sizes and their standard errors. A positive relationship between effect sizes and their standard errors has been termed ``small study effects'' (Sterne et al., 2000). The RoBMA-PSMA includes the precision-effect test (PET) and the precision-effect test with standard errors (PEESE) for small study effects. The PET model fits the relationship between observed effect sizes and their standard errors, while the PEESE model instead includes the square of the standard errors (i.e., their variances). The difference between the PET and PEESE models is that the PET model fits a linear relationship between the effect observed in a study and the precision with which that effect was estimated. The PEESE model fits a quadratic relationship. Thus, the PET and PEESE models differ with respect to the assumed underlying selection process. The PET model assumes that effect sizes increase in step with decreases in precision, consistent with selection based on statistical significance. The PEESE model assumes that studies with high precision are likely to be published regardless of statistical significance, whereas increasingly imprecise studies require increasingly larger effect size estimates to survive censorship.

@@

RoBMA-PSMA outperforms other meta-analytic approaches, including each constituent model included in RoBMA-PSMA, in simulations covering a range of plausible scenarios (Maier et al., in-press). Further, an analysis of real data from an approximately known data generating process (multi-lab registered replication report) suggested that RoBMA-PSMA outperforms other available techniques (Bartoš \& Schimmack, in-press). These strong performance indices combined with the ability to simultaneously model various plausible manifestations of reporting bias made RoBMA-PSMA an attractive choice for re-analyzing the enhanced expectancies and self-controlled practice meta-analyses.

@@

Similar to RoBMA-PSMA (Maier et al., in-press), \emph{z}-curve has also performed well in simulation studies and when applied to multi-lab registered replication data (Bartoš \& Schimmack, in-press). \emph{Z}-curve is a tool that can provide insight into the possible power shortcomings of a particular literature. Further, as \emph{z}-curve was designed to accommodate highly heterogeneous inputs, it is ideal for exploring power and reporting bias among studies that used a variety of methods to manipulate the two motivational factors in Wulf and Lewthwaite's (2016) OPTIMAL theory of motor learning.

@@

Statistical analyses were conducted using R (Version 4.2.1; R Core Team, 2021) and the R-packages \emph{geomtextpath} (Version 0.1.0; Cameron \& van den Brand, 2022), \emph{gt} (Version 0.6.0; Iannone et al., 2022), \emph{invgamma} (Version 1.1; Kahle \& Stamey, 2017), \emph{metafor} (Version 3.4.0; Viechtbauer, 2010), \emph{papaja} (Version 0.1.0.9999; Aust \& Barth, 2020), \emph{patchwork} (Version 1.1.0.9000; Pedersen, 2022), \emph{plotly} (Version 4.10.0; Sievert, 2020), \emph{pwr} (Version 1.3.0; Champely, 2020), \emph{renv} (Version 0.15.5; Ushey, 2022), \emph{robma} (Bartoš \& Maier, 2020), \emph{tidyverse} (Version 1.3.1; Wickham et al., 2019), \emph{tinylabels} (Version 0.2.3; Barth, 2022), \emph{truncdist} (Version 1.0.2; Novomestky \& Nadarajah, 2016), and \emph{zcurve} (Version 2.1.2; Bartoš \& Schimmack, 2020) were used in this project.

@@

The model-averaged posterior distribution of the average effect from the RoBMA-PSMA model is displayed in Figure \ref{fig:fig1}B. The results suggest moderate evidence against the presence of an effect, BF\textsubscript{01} = 3.16, very weak evidence against the presence of heterogeneity, BF\textsubscript{rf} = 1.7, and overwhelming evidence for the presence of reporting bias, BF\textsubscript{pb} = 18,399. The overall model ensemble estimated the effect of self-controlled practice as \emph{d} = .034 (95\% credible interval {[}.0, .248{]}). Heterogeneity was estimated as \(\tau\) = .05 (95\% credible interval {[}.0, .261{]}). A model fit with two influential cases included found overwhelming evidence for the presence of heterogeneity, BF\textsubscript{rf} = 1,924,516 and estimated \(\tau\) = .559 (95\% credible interval {[}.36, .78{]}). There were no other meaningful differences between models.

@@

The model-averaged posterior distribution of the average effect from the RoBMA-PSMA model is displayed in Figure \ref{fig:fig1}C. The results revealed weak evidence for the presence of an effect BF\textsubscript{10} = 1.9, weak evidence for the presence of reporting bias, BF\textsubscript{pb} = 2.3, and very strong evidence for the presence of heterogeneity, BF\textsubscript{rf} = 47.6. The model ensemble estimated an average effect of \emph{d} = .26 (95\% credible interval {[}-.07, .63{]}). Heterogeneity was estimated as \(\tau\) = .35 (95\% credible interval {[}.07, .54{]}). A model fit with two influential cases included found weak evidence for the absence of an effect BF\textsubscript{01} = 2.5, strong evidence for the presence of reporting bias, BF\textsubscript{pb} = 21, and overwhelming evidence for the presence of heterogeneity, BF\textsubscript{rf} = 45,300. The model ensemble estimated an average effect of \emph{d} = .00 (95\% credible interval {[}-.62, .59{]}), \(\tau\) = .49 (95\% credible interval {[}.32, .68{]}).

@@

\caption{\normalfont \small \textbf{Results from the robust Bayesian meta-analysis with publication selection model-averaging (RoBMA-PSMA) method.} \textbf{(A)} Prior spike and distribution (purple) with 50\% of prior probability density concentrated on the null hypothesis and 50\% reflecting plausible true effects with \emph{M} = 0 and \emph{SD} = 1. (The x-axis is truncated; the prior did include a small probability of effects larger than 1 and smaller than -1.) The model-averaged posterior distribution of the mean effect (Mu) of \textbf{(B)} self-controlled practice on motor learning (blue). The increased height of the spike at mu = 0 reflects increased belief in the null hypothesis. The remaining distribution reflects updated belief in the size and direction of a possible true effect. The mean estimate (\emph{M} = .034) for the posterior distribution is represented by the dashed vertical line. The model-averaged posterior distribution of the mean effect of \textbf{(C)} enhanced expectancies on motor learning (red). The decreased height of the spike at mu = 0 reflects decreased belief in the null hypothesis. The remaining distribution reflected updated belief in the size and direction of a possible true effect. The mean estimate (\emph{M} = .26) for the posterior distribution is represented by the dashed vertical line. The model-averaged posterior distribution generated from an analysis of \textbf{(D)} simulated data (green). In the simulation, 49 studies were sampled from a population with a true effect of mu = .54 and no reporting bias. The mean estimate (\emph{M} = .504) for the posterior is represented by the vertical dashed line. Outliers were not included in the self-controlled (\emph{n} = 2) and enhanced expectancies (\emph{n} = 2) models.}\label{fig:fig2}
\end{figure}

@@

\normalfont

@@

The results of the \emph{z}-curve analysis can be seen in Figure \ref{fig:fig2}C. The analysis estimated that the expected discovery rate of studies conducted on enhanced expectancies was 8\% (95\% confidence interval {[}5\%, 18\%{]}). The observed discovery rate was 44\% (95\% confidence interval {[}31\%, 59\%{]}). Since the upper bound of the expected discovery rate does not overlap with the lower bound of the observed discovery rate, there is evidence of significant reporting bias. The expected replication rate of the statistically significant results was 33\% (95\% confidence interval {[}8\%, 59\%{]}). Including influential cases did not meaningfully change the results.

@@

\caption{\normalfont \small \textbf{Results of the \emph{z}-curve analyses.} Distribution of \emph{z}-values for \textbf{(A)} our simulation with 47\% results and no reporting bias (green), \textbf{(B, left)} self-controlled practice (blue), \textbf{(C, left)} enhanced expectancies (red), and \textbf{(D, left)} motivational factors combined (purple). Values in the \emph{z}-score distributions for each analysis that are to the right of the significance line (\emph{z} = 1.96; solid, black) are statistically significant with a two-tailed \(\alpha\) of .05. Bootstrapped confidence distributions for the expected discovery rate (EDR; dark) and expected replication rate (ERR; light) for \textbf{(B, right)} self-controlled practice (blue), \textbf{(C, right)} enhanced expectancies (red), and \textbf{(D, right)} motivational factors combined (purple). Note, reported CIs include 5 extra points (EDR) and 3 extra points (ERR) added to the quantiles of the bootstrapped distributions in the right panel, consistent with Bartoš et al. (2022). The expected discovery rate is the estimated average power of all studies that have been conducted . The expected replication rate is the estimated power of all studies that observed a statistically significant result. The analysis estimated the average statistical power to be 6\%, 8\%, and 6\% for self-controlled practice, enhanced expectancies, and motivational factors combined, respectively. The estimated conditional power of the statistically significant results was 11\%, 33\%, and 21\% for self-controlled practice, enhanced expectancies, and motivational factors combined, respectively. Note that a minimum power of 80\% (dashed line, black) is often recommended. Outliers were not included in the self-controlled (\emph{n} = 2), enhanced expectancies (\emph{n} = 2), and motivational factors combined (\emph{n} = 4) analyses.}\label{fig:fig3}
\end{figure}

@@

Our objective in the current study was to conduct a holistic assessment of the motivation pillar of OPTIMAL theory. Specifically, we used RoBMA-PSMA---a state-of-the-art bias correction technique---to evaluate the effect of self-controlled practice and enhanced expectancies on motor learning (see Figure \ref{fig:fig1}). Additionally, we used \emph{z}-curve analyses to estimate the statistical power of the enhanced expectancies studies alone and then combined with the self-controlled practice studies to assess the motivation pillar of OPTIMAL theory (see Figure \ref{fig:fig2}). Using RoBMA-PSMA to model the effect of self-controlled practice on motor learning, we found moderate evidence against the presence of an effect (BF\textsubscript{01} = 3.16). The model ensemble estimated a small average effect of \emph{d} = .034 with a 95\% credible interval that reached zero {[}.0, .248{]}. We also found overwhelming evidence for the presence of reporting bias (BF\textsubscript{pb} = 18,399). When modeling the effect of enhanced expectancies on motor learning with RoBMA-PSMA, we found weak evidence for the presence of an effect (BF\textsubscript{10} = 1.9). The model ensemble estimated a small average effect of \emph{d} = .26 with a 95\% credible interval that included zero {[}-.07, .63{]}. Weak evidence for the presence of reporting bias (BF\textsubscript{pb} = 2.3) was also found. There was very strong support for the presence of heterogeneity in the enhanced expectancies literature (BF\textsubscript{rf} = 47.6).

Heterogeneity can be problematic for modeling reporting bias with the selection and regression models employed in RoBMA-PSMA (Carter et al., 2019). Fortunately, the \emph{z}-curve was designed for heterogeneous samples and may therefore be a better method of evaluating reporting bias in the enhanced expectancies data. The \emph{z}-curve analysis estimated the average statistical power of enhanced expectancies studies to be 8\%. The 95\% confidence interval {[}5\%, 18\%{]} did not overlap with the 95\% confidence interval of the observed discovery rate {[}31\%, 59\%{]}, providing evidence of significant reporting bias. Visual inspection of the distribution of \emph{z}-values reveals a large concentration of barely significant results and a dearth of barely not significant results (Figure \ref{fig:fig2}B). This pattern is consistent with selective reporting, and it is difficult to imagine an alternate process that would generate these results. To evaluate the potential presence of reporting bias across both motivation factors, as well as estimate the average power of studies that have investigated these phenomena, we applied a \emph{z}-curve to studies from both meta-analyses. The \emph{z}-curve analysis combining the enhanced expectancies and self-controlled practice studies estimated the average statistical power to be 6\%. The 95\% confidence interval {[}5\%, 13\%{]} did not overlap with the 95\% confidence interval of the observed discovery rate {[}37\%, 56\%{]}, indicating significant reporting bias.

@@

\leavevmode\vadjust pre{\hypertarget{ref-bartos2020}{}}%
Bartoš, F., \& Schimmack, U. (in-press). Z-curve 2.0: Estimating replication rates and discovery rates. \emph{Meta-Psychology}. \url{https://doi.org/10.31234/osf.io/urgtn}

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-mckay2022c}{}}%
McKay, B., Bacelar, M. F. B., \& Carter, M. J. (2022). \emph{\href{}{On the reproducibility of power analyses in motor behavior research}}. {SportRxiv}.

@@

>>>>

\noindent \textsuperscript{$\dagger$}Co-Senior Authors \newline
\newline
\noindent \addORCIDlink{Brad McKay}{0000-0002-7408-2323} \newline
\noindent \addORCIDlink{Mariane F.B. Bacelar}{0000-0002-0388-1769} \newline
\noindent \addORCIDlink{Juliana O. Parma}{0000-0002-3118-9921} \newline
\noindent \addORCIDlink{Matthew W. Miller}{0000-0002-7896-7217} \newline
\noindent \addORCIDlink{Michael J. Carter}{0000-0002-0675-4271} \newline       
\newline
\noindent Data and code: \url{https://osf.io/vfza7/} \newline
\newline
\noindent \textbf{Corresponding authors:} Brad McKay (\href{mailto:bradmckay8@gmail.com}{\nolinkurl{bradmckay8@gmail.com}}; \href{mailto:mckayb9@mcmaster.ca}{\nolinkurl{mckayb9@mcmaster.ca}}) and Michael J. Carter (\href{mailto:cartem11@mcmaster.ca}{\nolinkurl{cartem11@mcmaster.ca}}; \href{mailto:motorlab@mcmaster.ca}{\nolinkurl{motorlab@mcmaster.ca}})

@@

A challenge for most reporting bias models is large heterogeneity in true effects. While RoBMA appears to perform well with moderate levels of heterogeneity, its performance has not been evaluated when heterogeneity is high and the performance of each of its constituent reporting bias models suffers with high heterogeneity (Carter et al., 2019). The \emph{z}-curve model was designed specifically to perform well regardless of heterogeneity (Bartoš \& Schimmack, 2022). While \emph{z}-curve does not provide adjusted effect size estimates (and such estimates may be meaningless with high heterogeneity), it instead estimates the average underlying power of included experiments. A significant difference between the estimated power of studies and the observed proportion of significant results can indicate the presence of reporting bias in a literature.

Here, we leveraged state-of-the-art robust Bayesian meta-analysis and \emph{z}-curve methods to re-analyze the meta-analyses by McKay and colleagues (in-press) and Bacelar and colleagues (2022). Considering the potential importance of OPTIMAL theory for the field of motor learning and performance (see T. D. Lee \& Carnahan, 2021 for a discussion), it is imperative that the evidence buttressing its motivation predictions be evaluated as rigorously as possible. Critically, our re-analysis addresses limitations in both previous meta-analyses. First, the analysis of enhanced expectancies fit only one bias correction model---the trim-and-fill method (Duval \& Tweedie, 1998)---and that model has been shown to result in exaggerated effect size estimates and severely inflated Type 1 error rates in the presence of publication bias and small or null effects (Bartoš et al., 2022; Carter et al., 2019; Hong \& Reed, 2021). Second, although the results from multiple reporting bias models coalesced around small effect sizes (ranged from \emph{g} = -.11 to \emph{g} = .26) in the analysis of self-controlled practice, there are no principled reasons for preferring one estimate over another. We now consider a wider range of plausible models of reporting bias than that used in the previous meta-analyses. We also leverage Bayesian model averaging to upweight the best performing models, which has the advantage of evaluating single model-averaged posterior distributions for each parameter of interest. Lastly, we fit \emph{z}-curve models to the data in both meta-analyses. With this technique, the average power can be estimated and compared to the rate of significant results, providing crucial insight into the quality of the evidence-base supporting motivation predictions in OPTIMAL theory.

@@

The original meta-analyses followed similar inclusion criteria and data extraction policies (see Figure \ref{fig:fig1}). All data were extracted by two independent researchers with conflicts settled by including a third researcher. The literature search for the self-controlled practice meta-analysis (for further details see McKay et al., in-press) ended in August 2019 and in June 2020 for the enhanced expectancies meta-analysis (for further details see Bacelar, Parma, Murrah, et al., 2022).

@@

\caption{\normalfont \footnotesize \textbf{Description of the original meta-analyses.} Diagram outlining the shared and unique inclusion criteria, dependent variable selection, the number of studies screened, and the number of studies included in each meta-analysis.}\label{fig:fig1}
\end{figure}

@@

The second class of models fit in the RoBMA-PSMA method model the relationship between observed effect sizes and their standard errors. A positive relationship between effect sizes and their standard errors has been termed ``small study effects'' (Sterne et al., 2000). The RoBMA-PSMA includes the precision-effect test (PET) and the precision-effect test with standard errors (PEESE) for small study effects. The PET model fits the relationship between observed effect sizes and their standard errors, while the PEESE model instead includes the square of the standard errors (i.e., their variances) (Stanley \& Doucouliagos, 2014). The difference between the PET and PEESE models is that the PET model fits a linear relationship between the effect observed in a study and the precision with which that effect was estimated. The PEESE model fits a quadratic relationship. Thus, the PET and PEESE models differ with respect to the assumed underlying selection process (Stanley \& Doucouliagos, 2014). The PET model assumes that effect sizes increase in step with decreases in precision, consistent with selection based on statistical significance. The PEESE model assumes that studies with high precision are likely to be published regardless of statistical significance, whereas increasingly imprecise studies require increasingly larger effect size estimates to survive censorship.

@@

RoBMA-PSMA outperforms other meta-analytic approaches, including each constituent model included in RoBMA-PSMA, in simulations covering a range of plausible scenarios (Maier et al., in-press). Further, an analysis of real data from an approximately known data generating process (multi-lab registered replication report) suggested that RoBMA-PSMA outperforms other available techniques (Bartoš et al., 2022). These strong performance indices combined with the ability to simultaneously model various plausible manifestations of reporting bias made RoBMA-PSMA an attractive choice for re-analyzing the enhanced expectancies and self-controlled practice meta-analyses.

@@

Similar to RoBMA-PSMA (Maier et al., in-press), \emph{z}-curve has also performed well in simulation studies and when applied to multi-lab registered replication data (Bartoš \& Schimmack, 2022). \emph{Z}-curve is a tool that can provide insight into the possible power shortcomings of a particular literature. Further, as \emph{z}-curve was designed to accommodate highly heterogeneous inputs, it is ideal for exploring power and reporting bias among studies that used a variety of methods to manipulate the two motivational factors in Wulf and Lewthwaite's (2016) OPTIMAL theory of motor learning.

@@

Statistical analyses were conducted using R (Version 4.1.2; R Core Team, 2021) and the R-packages \emph{geomtextpath} (Version 0.1.0; Cameron \& van den Brand, 2022), \emph{gt} (Version 0.6.0; Iannone et al., 2022), \emph{invgamma} (Version 1.1; Kahle \& Stamey, 2017), \emph{metafor} (Version 3.4.0; Viechtbauer, 2010), \emph{papaja} (Version 0.1.0.9999; Aust \& Barth, 2020), \emph{patchwork} (Version 1.1.0.9000; Pedersen, 2022), \emph{plotly} (Version 4.10.0; Sievert, 2020), \emph{pwr} (Version 1.3.0; Champely, 2020), \emph{renv} (Version 0.15.5; Ushey, 2022), \emph{robma} (Bartoš \& Maier, 2020), \emph{tidyverse} (Version 1.3.1; Wickham et al., 2019), \emph{tinylabels} (Version 0.2.3; Barth, 2022), \emph{truncdist} (Version 1.0.2; Novomestky \& Nadarajah, 2016), and \emph{zcurve} (Version 2.1.2; Bartoš \& Schimmack, 2020) were used in this project.

@@

The model-averaged posterior distribution of the average effect from the RoBMA-PSMA model is displayed in Figure \ref{fig:fig2}B. The results suggest moderate evidence against the presence of an effect, BF\textsubscript{01} = 3.16, very weak evidence against the presence of heterogeneity, BF\textsubscript{rf} = 1.7, and overwhelming evidence for the presence of reporting bias, BF\textsubscript{pb} = 18,399. The overall model ensemble estimated the effect of self-controlled practice as \emph{d} = .034 (95\% credible interval {[}.0, .248{]}). Heterogeneity was estimated as \(\tau\) = .05 (95\% credible interval {[}.0, .261{]}). A model fit with two influential cases included found overwhelming evidence for the presence of heterogeneity, BF\textsubscript{rf} = 1,924,516 and estimated \(\tau\) = .559 (95\% credible interval {[}.36, .78{]}). There were no other meaningful differences between models.

@@

The model-averaged posterior distribution of the average effect from the RoBMA-PSMA model is displayed in Figure \ref{fig:fig2}C. The results revealed weak evidence for the presence of an effect BF\textsubscript{10} = 1.9, weak evidence for the presence of reporting bias, BF\textsubscript{pb} = 2.3, and very strong evidence for the presence of heterogeneity, BF\textsubscript{rf} = 47.6. The model ensemble estimated an average effect of \emph{d} = .26 (95\% credible interval {[}-.07, .63{]}). Heterogeneity was estimated as \(\tau\) = .35 (95\% credible interval {[}.07, .54{]}). A model fit with two influential cases included found weak evidence for the absence of an effect BF\textsubscript{01} = 2.5, strong evidence for the presence of reporting bias, BF\textsubscript{pb} = 21, and overwhelming evidence for the presence of heterogeneity, BF\textsubscript{rf} = 45,300. The model ensemble estimated an average effect of \emph{d} = .00 (95\% credible interval {[}-.62, .59{]}), \(\tau\) = .49 (95\% credible interval {[}.32, .68{]}).

@@

\caption{\normalfont \footnotesize \textbf{Results from the robust Bayesian meta-analysis with publication selection model-averaging (RoBMA-PSMA) method.} \textbf{(A)} Prior spike and distribution (purple) with 50\% of prior probability density concentrated on the null hypothesis and 50\% reflecting plausible true effects with \emph{M} = 0 and \emph{SD} = 1. (The x-axis is truncated; the prior did include a small probability of effects larger than 1 and smaller than -1.) The model-averaged posterior distribution of the mean effect (Mu) of \textbf{(B)} self-controlled practice on motor learning (blue). The increased height of the spike at mu = 0 reflects increased belief in the null hypothesis. The remaining distribution reflects updated belief in the size and direction of a possible true effect. The mean estimate (\emph{M} = .034) for the posterior distribution is represented by the dashed vertical line. The model-averaged posterior distribution of the mean effect of \textbf{(C)} enhanced expectancies on motor learning (red). The decreased height of the spike at mu = 0 reflects decreased belief in the null hypothesis. The remaining distribution reflected updated belief in the size and direction of a possible true effect. The mean estimate (\emph{M} = .26) for the posterior distribution is represented by the dashed vertical line. The model-averaged posterior distribution generated from an analysis of \textbf{(D)} simulated data (green). In the simulation, 49 studies were sampled from a population with a true effect of mu = .54 and no reporting bias. The mean estimate (\emph{M} = .504) for the posterior is represented by the vertical dashed line. Outliers were not included in the self-controlled (\emph{n} = 2) and enhanced expectancies (\emph{n} = 2) models.}\label{fig:fig2}
\end{figure}

@@

@@

The results of the \emph{z}-curve analysis can be seen in Figure \ref{fig:fig3}C. The analysis estimated that the expected discovery rate of studies conducted on enhanced expectancies was 8\% (95\% confidence interval {[}5\%, 18\%{]}). The observed discovery rate was 44\% (95\% confidence interval {[}31\%, 59\%{]}). Since the upper bound of the expected discovery rate does not overlap with the lower bound of the observed discovery rate, there is evidence of significant reporting bias. The expected replication rate of the statistically significant results was 33\% (95\% confidence interval {[}8\%, 59\%{]}). Including influential cases did not meaningfully change the results.

@@

\caption{\normalfont \footnotesize \textbf{Results of the \emph{z}-curve analyses.} Distribution of \emph{z}-values for \textbf{(A)} our simulation with 47\% results and no reporting bias (green), \textbf{(B, left)} self-controlled practice (blue), \textbf{(C, left)} enhanced expectancies (red), and \textbf{(D, left)} motivational factors combined (purple). Values in the \emph{z}-score distributions for each analysis that are to the right of the significance line (\emph{z} = 1.96; solid, black) are statistically significant with a two-tailed \(\alpha\) of .05. Bootstrapped confidence distributions for the expected discovery rate (EDR; dark) and expected replication rate (ERR; light) for \textbf{(B, right)} self-controlled practice (blue), \textbf{(C, right)} enhanced expectancies (red), and \textbf{(D, right)} motivational factors combined (purple). Note, reported CIs include 5 extra points (EDR) and 3 extra points (ERR) added to the quantiles of the bootstrapped distributions in the right panel, consistent with Bartoš et al. (2022). The expected discovery rate is the estimated average power of all studies that have been conducted . The expected replication rate is the estimated power of all studies that observed a statistically significant result. The analysis estimated the average statistical power to be 6\%, 8\%, and 6\% for self-controlled practice, enhanced expectancies, and motivational factors combined, respectively. The estimated conditional power of the statistically significant results was 11\%, 33\%, and 21\% for self-controlled practice, enhanced expectancies, and motivational factors combined, respectively. Note that a minimum power of 80\% (dashed line, black) is often recommended. Outliers were not included in the self-controlled (\emph{n} = 2), enhanced expectancies (\emph{n} = 2), and motivational factors combined (\emph{n} = 4) analyses.}\label{fig:fig3}
\end{figure}

@@

Our objective in the current study was to conduct a holistic assessment of the motivation pillar of OPTIMAL theory. Specifically, we used RoBMA-PSMA---a state-of-the-art bias correction technique---to evaluate the effect of self-controlled practice and enhanced expectancies on motor learning (see Figure \ref{fig:fig2}). Additionally, we used \emph{z}-curve analyses to estimate the statistical power of the enhanced expectancies studies alone and then combined with the self-controlled practice studies to assess the motivation pillar of OPTIMAL theory (see Figure \ref{fig:fig3}). Using RoBMA-PSMA to model the effect of self-controlled practice on motor learning, we found moderate evidence against the presence of an effect (BF\textsubscript{01} = 3.16). The model ensemble estimated a small average effect of \emph{d} = .034 with a 95\% credible interval that reached zero {[}.0, .248{]}. We also found overwhelming evidence for the presence of reporting bias (BF\textsubscript{pb} = 18,399). When modeling the effect of enhanced expectancies on motor learning with RoBMA-PSMA, we found weak evidence for the presence of an effect (BF\textsubscript{10} = 1.9). The model ensemble estimated a small average effect of \emph{d} = .26 with a 95\% credible interval that included zero {[}-.07, .63{]}. Weak evidence for the presence of reporting bias (BF\textsubscript{pb} = 2.3) was also found. There was very strong support for the presence of heterogeneity in the enhanced expectancies literature (BF\textsubscript{rf} = 47.6).

Heterogeneity can be problematic for modeling reporting bias with the selection and regression models employed in RoBMA-PSMA (Carter et al., 2019). Fortunately, the \emph{z}-curve was designed for heterogeneous samples and may therefore be a better method of evaluating reporting bias in the enhanced expectancies data. The \emph{z}-curve analysis estimated the average statistical power of enhanced expectancies studies to be 8\%. The 95\% confidence interval {[}5\%, 18\%{]} did not overlap with the 95\% confidence interval of the observed discovery rate {[}31\%, 59\%{]}, providing evidence of significant reporting bias. Visual inspection of the distribution of \emph{z}-values reveals a large concentration of barely significant results and a dearth of barely not significant results (Figure \ref{fig:fig3}B). This pattern is consistent with selective reporting, and it is difficult to imagine an alternate process that would generate these results. To evaluate the potential presence of reporting bias across both motivation factors, as well as estimate the average power of studies that have investigated these phenomena, we applied a \emph{z}-curve to studies from both meta-analyses. The \emph{z}-curve analysis combining the enhanced expectancies and self-controlled practice studies estimated the average statistical power to be 6\%. The 95\% confidence interval {[}5\%, 13\%{]} did not overlap with the 95\% confidence interval of the observed discovery rate {[}37\%, 56\%{]}, indicating significant reporting bias.

@@

\leavevmode\vadjust pre{\hypertarget{ref-bartos2020}{}}%
Bartoš, F., \& Schimmack, U. (2022). Z-curve 2.0: Estimating replication rates and discovery rates. \emph{Meta-Psychology}, \emph{6}. \url{https://doi.org/10.15626/MP.2021.2720}

@@

\leavevmode\vadjust pre{\hypertarget{ref-hong2021}{}}%
Hong, S., \& Reed, W. R. (2021). Using monte carlo experiments to select meta-analytic estimators. \emph{Research Synthesis Methods}, \emph{12}(2), 192--215.

@@

\leavevmode\vadjust pre{\hypertarget{ref-mckay2022c}{}}%
McKay, B., Bacelar, M. F. B., \& Carter, M. J. (2022). \emph{On the reproducibility of power analyses in motor behavior research}. {SportRxiv}. \url{https://doi.org/10.51224/SRXIV.184}

@@

\leavevmode\vadjust pre{\hypertarget{ref-stanley2014}{}}%
Stanley, T. D., \& Doucouliagos, H. (2014). Meta-regression approximations to reduce publication selection bias. \emph{Research Synthesis Methods}, \emph{5}(1), 60--78.
